### 2022-04-07 manuscript/header.tex
initial commit

\usepackage{amsmath}

% ---------- CHOOSE A FONT ----------

\usepackage{fontspec}
%\setmainfont{Fira Sans}
\setmainfont{Verdana}

\usepackage[english]{babel}							   % hyphenation etc for English

% --------- MISC FORMATTING ---------
%
\usepackage{setspace}                % needed for doublespacing
%\doublespacing                       % doublespaced line spacing
\usepackage{graphicx}                % for importing graphics into figures
\usepackage{float}                   % better control of figure placement
\usepackage{hyperref}                % for clickable URLs and email addresses
\usepackage{cleveref}
\urlstyle{sf}                        % san-serif font for URLs
\usepackage[margin=1.2in]{geometry}  % control page margins
\usepackage[labelfont=bf]{caption}   % make caption labels boldface
\usepackage[bottom]{footmisc}        % footnotes at bottom of page
\setlength{\skip\footins}{10mm}      % obsessing about footnote spacing
\setlength{\parskip}{1ex}            % space between paragraphs
%\setlength{\parindent}{3em}	     % paragraph indentation
\usepackage{authblk}                 % author and affiliation formatting
\renewcommand\Affilfont{\small}
\usepackage{booktabs}
\usepackage{pdflscape}

% Biblio
\usepackage[style=apa,
labeldate=year,
isbn=false,
eprint=false,
date=year,
url=false,
backend=biber]{biblatex}
\bibliography{references.bib}

%-----------
% Line numbers
\usepackage{lineno}                  % add line numbers to margin
\def\linenumberfont{\normalfont\footnotesize\sffamily} % line numbers
\setlength\linenumbersep{9mm}                          % line numbers
\linenumbers                                           % line numbers