### 2022-04-07 manuscript/ms_genomes.Rmd

---
title: "Draft genomes of the three northern hemisphere blue mussel species: *Mytilus edulis*, *Mytilus galloprovincialis* and *Mytilus trossulus*"
author:
  - Alexis Simon^1^*
  - Christine Arbiol^1^
  - Nicolas Bierne^1^
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_tex: yes
    number_sections: yes
    toc: no
    includes:
      in_header: header.tex
  html_document:
    toc: no
documentclass: article
# classoption: twocolumn, landscape, etc
papersize: a4
pagestyle: plain  # plain, empty, headings
geometry: margin=1in
#linestretch: 2
fontfamily: mathpazo
fontsize: 11pt
bibliography: 
  - references.bib
  - knitcitations.bib
# zotero: "My shared Zotero group"
csl: g3.csl
# csl: https://www.zotero.org/styles/ecology
link-citations: yes
linkcolor: RoyalBlue
urlcolor: RoyalBlue
links-as-notes: false
editor_options: 
  markdown: 
    wrap: sentence
---

<!--# Insert author affiliations below -->

\small

^1^ CNRS, Univ Montpellier, ...

`*` Corresponding author: [alexis.simon\@normalesup.org](mailto:alexis.simon@normalesup.org){.email}

\normalsize

<!--# Write Abstract and Keywords below -->

```{=tex}
\vspace{1cm}
\hrule
```
Using the 10X chromium long reads technology, we provide draft genomes for three closely related blue mussel species from the *Mytilus* species complex.
The objective was to produce affordable genomic resources for population and evolutionary genomic studies.
Genomes are fragmented but represent a large portion of the genome, with good sizes and BUSCO scores.

```{=tex}
\vspace{3mm}
\hrule
\vspace{5mm}
```
*Keywords*: *Mytilus edulis*, *Mytilus galloprovincialis*, *Mytilus trossulus*, Genome assembly, 10X chromium

```{=tex}
\bleft
\newpage
```
```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")
```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()
cite_options(citation_format = "pandoc")
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
```

<!--# Start writing your paper below -->

# Introduction

Write your introduction here.

------------------------------------------------------------------------

# Methods

## Biological material and DNA extraction

One individual for each species of interest was collected and processed fresh.

Collection locations: - *Mytilus galloprovincialis* Sète, France, - *M. edulis* ???
- *M. trossulus* ???

Whole mussels were placed in 50 mL falcon tubes containing 25 ml of TNES-Urea solution and incubated for 4-6 weeks at room temperature (TNES-Urea: 10 mM Tris-HCl pH 7.4; 120 mM NaCl; 10 mM EDTA pH 8.0; 0.5% SDS; 4 M urea).

After this period of pre-treatment at ambient temperature, proteinase K was added at a final concentration of 150 $\mu$g/mL and the solution was incubated overnight at 56°C.

High Molecular weight genomic DNA was extracted following (From Nakayama et al., 1994, Chromosoma).
We used a 15 mL Phase Lock Gel Heavy extraction with 3 steps of phenol/chloroform/isoamylalcool (25/24/1) Tris pH 8,1 followed by 2 chloroform extractions.
After the last extraction, the aqueous supernatant was precipitated with 2 volumes of 100% EtOH and the pellet was hooked from the solution with a sterile glass Pasteur pipette.
The pellet was rinsed several times in 80% EtOH before being dried at room temperature.

DNA was resuspended with an appropriate volume of biomolecular water at 65°C for several hours, the duration of this incubation depending on the size of the granules.
DNA was then stored at 4°C.

Prior to the construction of the DNA libraries, DNA was repaired with NEBNext FFPE DNA Repair Mix according to the manufacturer's instructions.

## Library preparation and sequencing

The 10X chromium library preparation and sequencing was subcontracted to the MGX platform (Montpellier, France).

They followed 10X Genomics Genome Reagen Kit (*Genome Solution*) protocol to produce the 10X linked reads library using a Chromium microfluidic chip.

The library was sequenced on an Illumina NovaSeq 6000 with an S4 flowcell.

## Draft genome assemblies

Explain here v1, v4, v5

-   v1: `Supernova` v2.1.1
-   v4: `purge_dups` (`longranger align`)
-   v5: `agouti` using RNAseq data (`augustus` to get gff3, RNAseq `rcorrector` + `trimgalore` + `bwa-mem2` + merge bams). RNAseq data are from both NCBI and Maurine's/Erika's experiments.
-   v6: blobtoolkit filtration based on groups

-\> gives mtro_01, medu_01, mgal_01

## *M. trossulus* assembly improvement using minion reads and *M. coruscus* assembly

-   `LRScaf` using ONT reads (`minimap2` alignment)
-   `RagTag` using *M. coruscus* chromosomal assembly
-   `Pilon` using PE reads from 10X library (debarcoded + filtered with `fastp`, `bwa-mem2` alignement) + ONT reads

-\> gives mtro_02 (tros_v7)

## Repeats ???

## Annotation ???

## Document softwares, versions and DOIs:

-   Supernova v2.1.1 [cite](https://doi.org/10.1101/gr.214874.116)
-   Longranger v2.2.2
-   proc10xG <https://github.com/ucdavis-bioinformatics/proc10xG> commit 7afbfcf
-   nubeam-dedup <https://github.com/daihang16/nubeamdedup> commit 25dd385 [cite](https://doi.org/10.1093/bioinformatics/btaa112)
-   KMC <https://github.com/tbenavi1/KMC> commit 1df71f6
-   GenomeScope 2.0 <https://github.com/tbenavi1/genomescope2.0> commit 5034ed4 [cite](https://doi.org/10.1038/s41467-020-14998-3)
-   (? Smudgeplot commit 950d7b5)
-   Purge_Dups <https://github.com/dfguan/purge_dups> commit e1934bb [cite](https://doi.org/10.1093/bioinformatics/btaa025)
-   AGOUTI <https://github.com/svm-zhang/AGOUTI> commit a7e65d6 [cite](https://doi.org/10.1186/s13742-016-0136-3)
-   augustus v3.3.3 (for AGOUTI analysis) [@Stanke2008]
-   Blobtoolkit v2.4.0 [cite](https://doi.org/10.1534/g3.119.400908)
-   KAT v2.4.2 [cite](https://doi.org/10.1093/bioinformatics/btw663)
-   fastp v0.20.1 [@Chen2018a]
-   samtools v1.12 [@Li2009]
-   minimap2 v2.17 [@LiLi2017]
-   bwa-mem2 v2.2.1 [@VasimuddinMd2019]
-   bedtools v2.29.2 [@Quinlan2010]
-   seqkit v0.13.2 [@Shen2016]
-   rcorrector v1.0.4
-   trim-galore v0.6.6 [cite](https://github.com/FelixKrueger/TrimGalore) (Cutadapt + FastQC wrapper)
-   busco v5.1.1 [@Seppey2019]
-   assembly_stats v0.1.4 [cite](https://doi.org/10.5281/zenodo.3968774)
-   ragtag v1.1.1 [@Alonge2019]
-   pilon v1.24 [@Walker2014]
-   TETools DFAM container v1.3.1
-   RepeatModeler v2.0.1 [@Flynn2020]
-   cd-hit v4.8.1 [@Fu2012]
-   RepeatMasker v4.1.2-p1 (Smit, AFA, Hubley, R. RepeatModeler Open-1.0. 2008-2015 <http://www.repeatmasker.org>) [@Smit2013]

------------------------------------------------------------------------

# Results and discussion

Trees in forest *A* grew taller than those in forest *B* (mean height: `r mean(25, 31, 28)` versus `r mean(13, 19, 16)` m).

And many more cool results that get updated dynamically, e.g. see Table \@ref(tab:Table-mtcars) and Fig.
\@ref(fig:scatterplot).
Note Tables and Figures are **cross-linked and numbered automatically**.
They could also appear in the middle of the document, not necessarily at the end.

See also Fig.
\@ref(fig:SupplFig) and Table \@ref(tab:SupplTable) in the Supplementary Material.

```{r}
# asm_show <- c(
#   "coruscus_GCA017311375", 
#   "gallo_GCA900618805",
#   "edu_GCA905397895",
#   "eduam_GCA019925415",
#   "gallo_v6",
#   "gallo_v7",
#   "edu_v6", 
#   "edu_v7",
#   "tros_v6", 
#   "tros_v7"
# )
```

## Assembly statistics

```{r, include=FALSE}
source("../scripts/asm_stats_prep.R", local = knitr::knit_global())
```

```{r table_asm_stats}
asm_stats_table %>%
  kable(.)
```

## BUSCO scores

```{r load_busco}
busco <- read_tsv("data/busco_summary.tsv", col_types = cols()) %>% 
  mutate(asm = str_replace(asm, "_metazoa_odb10|_mollusca_odb10", "")) %>% # clean asm names
  mutate(cat = fct_relevel(cat, "M", "F", "CD", "CS")) %>%
  mutate(asm = fct_relevel(asm, asm_show) %>% 
           fct_rev()) %>% 
  group_by(asm, db) %>% 
  mutate(proportion = N/cur_data()$N[cur_data()$cat=="T"]) %>% 
  mutate(pct_fmt = paste(format(round(proportion*100, digits = 2)), "%"))
```

```{r full_figure_mollusca, fig.width=10, fig.height=4}
busco %>% 
  filter(db == "mollusca_odb10") %>% 
  filter(cat %in% c("CS", "CD", "F", "M")) %>%
  filter(asm %in% asm_show) %>% 
  ggplot(., aes(x = N, y = asm, fill = cat)) + 
    geom_bar(position = "fill", stat = "identity", width = 0.9) +
    scale_fill_manual(values = c("#D55E00", "#E69F00", "#56B4E9", "#009E73"),
                      guide = guide_legend(reverse=T),
                      name = "categories", 
                      labels = c("Missing", "Fragmented", "Complete duplicated", "Complete single copy")) +
    scale_x_continuous(expand=c(0,0), labels = scales::percent, breaks = c(0,0.2,0.4,0.6,0.8,1)) + 
    scale_y_discrete(expand=c(0,0)) +
    geom_text(aes(label = N), position = position_fill(vjust = .5), size = 3) +
    labs(x = "BUSCO percentage", y = NULL, title = "BUSCO scores Mollusca_odb10 (5295 genes)") +
    theme_minimal() +
    theme(axis.ticks.x = element_line(),
          panel.grid = element_blank(),
          legend.position = "bottom",
          plot.margin = margin(t=5,r=15,b=0,l=5))
```

```{r full_figure_metazoa, fig.width=10, fig.height=4}
busco %>% 
  filter(db == "metazoa_odb10") %>% 
  filter(cat %in% c("CS", "CD", "F", "M")) %>%
  filter(asm %in% asm_show) %>% 
  ggplot(., aes(x = N, y = asm, fill = cat)) + 
    geom_bar(position = "fill", stat = "identity") +
    scale_fill_manual(values = c("#D55E00", "#E69F00", "#56B4E9", "#009E73"),
                      guide = guide_legend(reverse=T),
                      name = "categories", 
                      labels = c("Missing", "Fragmented", "Complete duplicated", "Complete single copy")) +
    scale_x_continuous(expand=c(0,0), labels = scales::percent, breaks = c(0,0.2,0.4,0.6,0.8,1)) + 
    scale_y_discrete(expand=c(0,0)) +
    geom_text(aes(label = N), position = position_fill(vjust = .5), size = 3) +
    labs(x = "BUSCO percentage", y = NULL, title = "BUSCO scores Metazoa_odb10 (954 genes)") +
    theme_minimal() +
    theme(axis.ticks.x = element_line(),
          panel.grid = element_blank(),
          legend.position = "bottom",
          plot.margin = margin(t=5,r=15,b=0,l=5))
```

# Conclusion

Wrap up

# Acknowledgement

On the shoulders of giants.

# References

```{r write_citations, cache=FALSE, include=FALSE}
#write.bibtex(file = "knitcitations.bib")
```

::: {#refs}
:::

\eleft

\clearpage

<!--# Tables below -->

\listoftables

\newpage

```{r Table-Iris}
kable(head(iris), caption = "A glimpse of the famous Iris dataset.", booktabs = TRUE)
```

\clearpage

```{r Table-mtcars}
kable(mtcars[10:16, ], caption = "Now a subset of mtcars dataset.", booktabs = TRUE)
```

\clearpage

<!--# Figures below -->

\listoffigures

\newpage

```{r scatterplot, fig.cap="Just my first figure with a very fantastic caption."}
x <- rnorm(100)
y <- jitter(x, 1000)
plot(x, y)
```

\newpage

\blandscape

```{r Fig-landscape, fig.cap="Second figure in landscape format."}
a <- sort(rnorm(100))
b <- c(rep("Group Small", 35), rep("Group Big", 65))
boxplot(a ~ b)
```

\elandscape

\clearpage

# SUPPLEMENTARY MATERIAL {.unnumbered}

\beginsupplement

# Other things we tried

## Raw reads filtration

Explain pipeline, it did not improve assembly -\> v2, v3

\clearpage

```{r SupplTable}
kable(head(iris), caption = "A supplementary table.", booktabs = TRUE)
```

\clearpage

```{r SupplFig, fig.cap="A Supplementary Figure."}
x <- rnorm(100)
y <- jitter(x, 1000)
plot(x, y)
```

\newpage

```{r sessioninfo, eval = FALSE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
# sessioninfo::session_info()
```